---
title: "Lab3"
author: "Hugo Sparrman"
format: html
editor: visual
execute: 
  echo: false
---

```{r, message = FALSE}
library(here)
here::i_am("Lab3.qmd")
```

```{r, message = FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(stringr)
```

## Excercise 1

```{r}
spotify<-readRDS("Data/top-spotify-hits-2000-2019.Rds")
```

```{r}
unique_songs<-spotify|>
  distinct(song, artist, keep_all = TRUE)

```

### Question 1

The data set contains: `r nrow(unique_songs)` different songs and `r nrow(unique_songs|>distinct(artist))` distinct number of artists, as well as `r nrow(spotify|>distinct(genre))` distinct number of genres.

### Question 2

```{r}
artist_by_year<-spotify|>
  group_by(year)|>
  distinct(artist)|>
  summarise(number_of_artists=n())

knitr::kable(artist_by_year,caption = "Distinct number of artists by year",
             col.names = c("Year","Number of Artists"))
```

### Question 3

```{r}
popular_song <-spotify|>
  distinct(artist,song)|>
  group_by(artist)|>
  summarise(Number_of_songs=n())|>
  arrange(-Number_of_songs)|>
  slice(1)

```

The most popular artist in the data set is: `r popular_song$artist` with `r popular_song$Number_of_songs` number of songs in the data set.

### Question 4

```{r}
genre_table <- spotify|>
  distinct(artist,song,.keep_all = TRUE)|>
  group_by(genre)|>
  summarise(
    min_tempo = min(tempo),
    max_tempo = max(tempo),
    mean_tempo = mean(tempo),
    median_tempo = median(tempo),
    number_of_songs = n_distinct(song)
  )

knitr::kable(genre_table,
             col.names = gsub("[_]"," ",str_to_title(names(genre_table))))

```

### Question 5

Creation of a data frame for the median of liveness and danceability. The result here is the five first rows.
```{r}
median_of_tings <- spotify|>
  group_by(year)|>
  summarise(
    nrgy = median(liveness),
    dnce = median(danceability)
  )
median_of_tings <-as.data.frame(median_of_tings)

head(median_of_tings,5)

```

### Question 6

```{r}
longer_median<-median_of_tings|>
  pivot_longer(cols = c(nrgy, dnce),
               names_to = "Variable",
               values_to = "Median")

ggplot(longer_median, aes(x = year, y = Median, color = Variable)) +
  geom_line() +
  labs(x = "Year", y = "Median", color = "Graphs") +
  ggtitle("Temporal evolution of median danceability and median liveness")


```

## Excercise 2

```{r}
dropout<-readRDS("Data/dropout.Rds")
```

### Question 1

We change the data frame dropout data to get conditional median age at enrollment. The output result can be seen in question 2.  
```{r}
age_at_enrollment<-dropout|>
  group_by(`Marital status`,`Gender`)|>
  summarise("Median age at enrollment" = median(`Age at enrollment`),.groups = "drop")
```

### Question 2

```{r}
Marital_status<-age_at_enrollment|>
  pivot_wider(names_from = Gender, values_from = `Median age at enrollment` )

knitr::kable(Marital_status, caption = "Median age of Marital situation ")

```

### Question 3

Here a data frame is created of the mean for all variables related to curricular units given the value of the "Target" variable. 
```{r}
mean_curricular<-dropout|>
  group_by(Target)|>
  summarise(across(starts_with("Curricular"),~mean(.,na.rm = T)))

```

### Question 4

```{r}
mean_table<-mean_curricular|>
  pivot_longer(cols = starts_with("Curricular"), names_to = "Units",
               values_to = "Value")|>
  pivot_wider(names_from = Target, values_from = "Value")

mean_table|>
  knitr::kable("pipe",caption = "Mean Curricular Units")
  
```
